Quotient maps and configuration spaces of hard disks

### 2022-02-10 manuscript/main.tex
eq 4 updated
proof zk is equal to z-k is added

All other quotient spaces are initially embedded in an infinite-dimensional descriptor space, and a numerical approach is used to estimate the minimum number of descriptors necessary to maintain the embedding. \red{This idea is analogous to representing a function in the Fourier space, but using a finite subset of infinitely-many coefficients.} Given a configuration space of $n$ disks, the distribution $f$ is defined as a sum of Dirac-delta distributions $\delta(\vec{a}_j)$ located at the disk centers $\vec{a}_j$ in the $a_1a_2$-coordinate system in Fig.\ \ref{fig:figure2}, or
\begin{equation}
f(\vec{a}) = \sum_{j = 1}^{n} \delta(\vec{a}_j) = \sum_{\vec{k}} {c_{\vec{k}} \exp{(2\pi \rmi \vec{k} \cdot \vec{a})}}
\label{eq:f}
\end{equation}
where $c_{\vec{k}}$ are the complex coefficients of the reciprocal lattice expansion and $\vec{k} = [p, q]$ for integers $p$ and $q$. The infinite set of $c_{\vec{k}}$ can be calculated using the orthogonality of the complex exponentials as
\begin{equation}
c_{\vec{k}} = \sum_{j=1}^{n} \exp{(-2\pi \rmi \vec{k} \cdot \vec{a}_j)} .
\label{eq:c_k}
\end{equation}
$c_{\vec{k}}$ respects the periodicity of the lattice and is invariant to permutations of the disk labels due to the commutative property of the summation in Eq.\ \ref{eq:f}. It can be shown that translating a configuration (by adding an offset to the $\vec{a}_j$) only changes the phase of the coefficients. This means that the moduli of the coefficients, or
\begin{equation}
z_{\vec{k}} = \sqrt{c^\ast_{\vec{k}} c_{\vec{k}}}
\label{eq:z_k}
\end{equation}
where $^\ast$ denotes the complex conjugate, are a set of real-valued descriptors that are invariant to disk label permutations and rigid translations. Observe that the descriptors $z_{\vec{k}}$ also respect inversion symmetry. Numerical experiments indicate that the rank of the Jacobian of the map from the $\vec{a}_j$ to the $z_{\vec{k}}$ is generically $2(n - 1)$, suggesting that some number of these descriptors could be sufficient to construct an embedding of $\Lambda / \mathcal{S}_2$.

@@

>>>>

All other quotient spaces are initially embedded in an infinite-dimensional descriptor space, and a numerical approach is used to estimate the minimum number of descriptors necessary to maintain the embedding. \red{This idea is analogous to representing a function in the Fourier space, but using a finite subset of infinitely-many coefficients.} Given a configuration space of $n$ disks, the distribution $f$ is defined as a sum of Dirac-delta distributions $\delta(\vec{a}_j)$ located at the disk centers $\vec{a}_j$ in the $a_1a_2$-coordinate system in Fig.\ \ref{fig:figure2}, or
\begin{equation}
	f(\vec{a}) 	= \sum_{j = 1}^{n} \delta(\vec{a}_j) = \sum_{\vec{k}} {c_{\vec{k}} \exp{(2\pi \rmi \vec{k} \cdot \vec{a})}}
	\label{eq:f}
\end{equation}
where $c_{\vec{k}}$ are the complex coefficients of the reciprocal lattice expansion and $\vec{k} = [p, q]$ for integers $p$ and $q$. The infinite set of $c_{\vec{k}}$ can be calculated using the orthogonality of the complex exponentials as
\begin{equation}
c_{\vec{k}} = \sum_{j=1}^{n} \exp{(-2\pi \rmi \vec{k} \cdot \vec{a}_j)} .
\label{eq:c_k}
\end{equation}
$c_{\vec{k}}$ respects the periodicity of the lattice and is invariant to permutations of the disk labels due to the commutative property of the summation in Eq.\ \ref{eq:f}. It can be shown that translating a configuration (by adding an offset to the $\vec{a}_j$) only changes the phase of the coefficients. This means that the moduli of the coefficients, or
\begin{equation}
z_{\vec{k}} = \sqrt{c^\ast_{\vec{k}} c_{\vec{k}}}
\label{eq:z_k}
\end{equation}
where $^\ast$ denotes the complex conjugate, are a set of real-valued descriptors that are invariant to disk label permutations and rigid translations. Observe that the descriptors $z_{\vec{k}}$ also respect inversion symmetry. \red{An illustration of the procedure above is provided in Fig.\ \ref{fig:app1} in App.\ \ref{sec:descriptor_proofs}. } Numerical experiments indicate that the rank of the Jacobian of the map from the $\vec{a}_j$ to the $z_{\vec{k}}$ is generically $2(n - 1)$, suggesting that some number of these descriptors could be sufficient to construct an embedding of $\Lambda / \mathcal{S}_2$.

@@

	This section provides proofs and supporting materials to show the relations among the descriptors. Let $\vec{x}$ and $\vec{a}$ respectively be the coordinates of a vector in the $xy$-coordinate system and $a_1a_2$-coordinate system. Let $\mat{T}$ be the forward coordinate transformation matrix such that $\vec{a} = \mat{T} \vec{x}$ and $\vec{x} = \mat{T}^{-1} \vec{a}$.

	The procedure to calculate the descriptors $z_{\vec{k}}$ is illustrated in Fig.\ \ref{fig:app1}. For a configuration $\config{p}$ whose coordinates are given in $xy$-coordinate system, the first step is to find the corresponding coordinates of the disks in the $a_1a_2$-coordinate system, and define the function $f(\vec{a})$ using Eq.\ \ref{eq:f}. Observe that the resulting function will be the same for each $n!$ permutations of the disk labels. Then, the complex coefficients of the infinite expansion in Eq.\ \ref{eq:f} are calculated using the orthogonality of complex exponentials as shown in Eq.\ \ref{eq:c_k}. The final step is then to take the modulus of these complex coefficients using Eq.\ \ref{eq:z_k}. Observe that the complex coefficients and the descriptors are sorted with respect to the distance of $\vec{k}$ from the origin. Moreover, observe that the complex coefficients $c_{\vec{k}}$ and $c_{-\vec{k}}$ have the same descriptors $z_{\vec{k}}$ for a given configuration.

	\begin{figure}
		\centering
		\includegraphics[width=1.0\columnwidth]{app1.eps}
		\caption{$3!$ permutations of a given configuration will result in the same function $f(\vec{a})$ after using Eq.\ \ref{eq:f}. The complex coefficients are calculated using Eq.\ \ref{eq:c_k}. Finally, the descriptors $z_{\vec{k}}$ are then defined as the modulus of the complex coefficients.}
		\label{fig:app1}
	\end{figure}

	The complex coefficients $c_{\vec{k}}$ is equal to $c^{\ast}_{-\vec{k}}$ where $^{\ast}$ denotes complex conjugate.
	\begin{proof}
		Let $\config{p}$ be a configuration and $\vec{p}$ be the coordinates of the disks in $a_1a_2$-coordinate system.
		\begin{align*} 
		c_{\vec{k}}(\config{p}) &=  \sum_{j=1}^{n} \exp{(-2\pi \rmi \vec{k} \cdot \vec{p}_j)} \nonumber \\ 
		&= \left\{ \sum_{j=1}^{n} \exp{(2\pi \rmi \vec{k} \cdot \vec{p}_j)} \right\}^{\ast} \nonumber \\
		&= \left\{ \sum_{j=1}^{n} \exp{[-2\pi \rmi (-\vec{k} \cdot \vec{p}_j)]} \right\}^{\ast} \nonumber \\
		&= c_{-\vec{k}}^{\ast}(\config{p})
		\end{align*}
	\end{proof}

	The descriptors $z_{\vec{k}}(\config{p}) = z_{-\vec{k}}(\config{p})$.
	\begin{proof}
		Using the fact that $c_{\vec{k}} = c^{\ast}_{-\vec{k}}$
		\begin{align*} 
		z_{\vec{k}}(\config{p}) =  \sqrt{c_{\vec{k}}(\config{p}) c_{\vec{k}}^\ast(\config{p})} = \sqrt{c^{\ast}_{-\vec{k}}(\config{p}) c_{-\vec{k}}(\config{p})} = z_{-\vec{k}}(\config{p})
		\end{align*}
	\end{proof}
